---
title: "Lab 3: Generalized Linear Models"
editor: source
date: 10-26-2022
draft: false
---

## R Packages Used

```{r}
library(VGAM) # Multinomial and Ordinal Regression
library(survival) # Data sets
library(MASS) # Data Set
library(GLMsData) # Data Set
library(tidyverse)
library(haven) # Reading Stata File
```

## References

For more information on this topic, I recommend you reading the following books:

1. Generalized Linear Models with Examples in R - Dunn

2. Vectorized Generalized Linear and Additive Models - Yee

## Generalized Linear Models

In linear regression, we are fitting a model between a set of predictors and an outcome variable. The outcome variable is generally normally distributed. However, if the outcome is not normally distributed, you must use a generalized linear model to assess the association between a set of predictors and an outcome variable.

### Link Functions

Due to the outcome variable being non-normally distributed, we must use link functions. The link function allows us to link the outcome variable to a linear model.

### Logistic Regression

A logistic regression model is used when the outcome is binary, yes and no. The link function is the logit function that models the association between a set of linear predictors and the odds of observing yes:

$$
\log\left\{\frac{P(Y=1)}{P(Y=0)}\right\} = \boldsymbol{X^T\beta}
$$
Fitting a logistic regression requires you to use the `glm()` function and specifying `family=binomial`. Then use the summary function to obtain the coefficients and `broom::tidy()`^[The double colon tells R to use a specific function from a package. Here, use the tidy function from the broom package.]

```{r}
logit_glm <- glm(y ~ ap + hilo, data = bacteria, 
                 family = binomial)
summary(logit_glm)

## OR ##

logit_glm %>% broom::tidy(exp=T)
```

A $\beta$ coefficient is interpreted as an odds ratio. The odds of success when `ap=p` is `r round(exp(0.7933),2)` times higher than when `ap=a`, after adjusting for `hilo`.


### Poisson Regression

Poisson regression is used when the outcome is count data. The link function is the log of the mean count:

$$
\log(\mu)=\boldsymbol{X^T\beta}
$$

Fitting a Poisson regression requires you to use the `glm()` function and specifying `family=poisson`. Then use the summary function to obtain the coefficients and `broom::tidy()`^[The double colon tells R to use a specific function from a package. Here, use the tidy function from the broom package.]

```{r}
pois_glm <- glm(recur ~ treatment + number, data = bladder1, 
                family = poisson)
summary(pois_glm)
pois_glm %>% broom::tidy(exp=T)
```
The line `pois_glm %>% broom::tidy(exp=T)` provides the exponentiated values of the $\beta$ coefficients. Interpreting the `number` variable, as the initial number of tumors increases, the mean number of recurrences increases by a factor of 1.12, adjusting for treatment status.


### Multinomial Regression

Multinomial Regression is used when the outcome is categorical variable with 3 or more categories. We fit a model using a logit link function for each category and the reference variable. For $k=1,\ldots,m$ categories and reference level $2$, the logit link function is modeled as:

$$
\log\left\{\frac{P(Y=j)}{P(Y=2)}\right\}=\boldsymbol{X^T\beta}\ \ j\in\{1,3,4,5,\ldots,m\}
$$

You can fit a multinomial regression model when using the `vglm()` function and setting `family = multinomial(refLevel=k)`, where `k` is your reference group. For more information of the data, visit this site: https://stats.oarc.ucla.edu/r/dae/multinomial-logistic-regression/.


```{r}
ml <- read_dta("https://stats.idre.ucla.edu/stat/data/hsbdemo.dta")
mglm <- vglm(factor(prog) ~ factor(ses) + write, data = ml, family = multinomial(refLevel = "2"))
summary(mglm)
exp(coef(mglm))
```

The line `exp(coef(mglm))` provides the exponentiated `\beta` coefficients for categorical variable. Interpreting `write` for `group=1` (academic): as writing score increased by 1 unit, the odds of of being in the academic group decreases by a factor of 0.94, adjusting of social economic status. 


### Ordinal Regression

Ordinal regression is a subclass of multinomial regression, but the categories have an ordinal component to it. For example, a likert scale can be considered ordinal: Strongly Disagree to Strongly Agree. An ordinal regression model uses the logit link function for model the probability of observing something greater:

$$
\log\left\{\frac{P(Y\ge j)}{P(Y<j)}\right\}=\boldsymbol{X^T\beta}\ \ j\in\{1,2,\ldots,m\}
$$
You can fit a ordinal regression model when using the `vglm()` function and setting `family = propodds`.


```{r}
dat <- read_dta("https://stats.idre.ucla.edu/stat/data/ologit.dta")
oglm <- vglm(apply~pared+public+gpa, data = dat, family = propodds)
summary(oglm)
exp(coef(oglm))
```
Intepreting `gpa`, as gpa increases by 1 point, the odds of being likely to apply increases by a factor of 1.851, adjusting for parental undergraduate education and plubic vs private college. 

## Questions

1.  What is the link function for a gamma distribution?

2. When will you use gamma regression?

3. The `lime` data set contains 385 observations on small-leaved lime trees. The `Foliage` variable measures the foliage biomass. Fit a regression model between `Foliage` and the following covariates: `DBH` tree diameter and `Age`. Make sure to print out your results.

**Submit your assignment as an html file. You can use either a QMD or RMD file to create the html**
