---
title: "Central Limit Theorem"
editor: visual
date: 10-23-2022
draft: false
---

## Central Limit Theorem

Let $X_1, X_2, \ldots, X_n$ be identical and independent distributed random variables with $E(X_i)=\mu$ and $Var(X_i) = \sigma²$. We define

$$
Y_n = \sqrt n \left(\frac{\bar X-\mu}{\sigma}\right) \mathrm{ where }\ \bar X = \frac{1}{n}\sum^n_{i=1}X_i.
$$

Then, the distribution of the function $Y_n$ converges to a standard normal distribution function as $n\rightarrow \infty$. The central limit theorem can b reexpressed as:

$$
\bar X \sim N(\mu,\sigma^2/n)
$$

## Normal Assumption

To check whether the data follows a normal distribution, we can utilize a [QQ-Plot](https://en.wikipedia.org/wiki/Q%E2%80%93Q_plot). The main idea is that the points, or quantiles must follow $f(x)=x$. To create a qq-plot in R, you will use `qqnorm()`. This will plot the points; additionally, you can use `qqline()` function to add $f(x)=x$ to the plot[^1] The example below will generate 1000 random variables from a standard normal distribution. Then it will create the qq-plot for you to evaluate. You will need to run `qqnorm` and `qqline` the same time to get the plot.

[^1]: You will need to run both of these lines at the same time.

```{r}
z_norm <- rnorm(1000)
qqnorm(z_norm)
qqline(z_norm)
```

The closer all the points are to the line, the more proof you have that the sample came from a normal distribution.

## Normal Distribution

The normal distribution has 2 finite moments. While the distribution of $\bar X$ will be normal regardless of sample size, the central limit theorem will still apply.

The code below provides a small simulation study where we will generate `nreals` random variables from standard normal distribution `nsim` times. Then we obtain the sample mean for each simulated data set and plot the histogram and QQ-plot.

```{r}
nsims <- 1000
nreals <- 10000
x <- matrix(nrow = nreals, ncol = nsims)
for(i in 1:nsims){
  x[,i]<- rnorm(nreals)
}
xbar <- colMeans(x)
plot(density(xbar))
qqnorm(xbar)
qqline(xbar)
```

## Cauchy Distribution

The [Cauchy Distribution](https://en.wikipedia.org/wiki/Cauchy_distribution) has and undefined mean and variance; therefore, the central limit theorem does not apply to it.

```{r}
nsims <- 1000
nreals <- 10000
x <- matrix(nrow = nreals, ncol = nsims)
for(i in 1:nsims){
  x[,i]<- rcauchy(nreals)
}
xbar <- colMeans(x)
plot(density(xbar))
qqnorm(xbar)
qqline(xbar)
```

## Problems

### Problem 1

Comment the code below describing the what each line is doing:

```{r}
nsims <- 1000
nreals <- 10000
x <- matrix(nrow = nreals, ncol = nsims)
for(i in 1:nsims){
  x[,i]<- rnorm(nreals)
}
xbar <- colMeans(x)
plot(density(xbar))
qqnorm(xbar)
qqline(xbar)
```

## Problem 2

Let $X_1,\ldots, X_n\overset{iid}{\sim}Bernoulli(0.4)$. Does $\bar X$ follow a normal distribution when the sample size gets larger.

### Problem 3

Let $X_1,\ldots, X_n\overset{iid}{\sim}Gamma(3,4)$. Does $\bar X$ follow a normal distribution when the sample size gets larger.

### Problem 3

Let $X_1,\ldots, X_n\overset{iid}{\sim}Beta(1,2)$. Does $\bar X$ follow a normal distribution when the sample size gets larger.

**Complete the assignment and submit your code as a QMD file. Submit your file to Canvas on 11/2/2022 at 11:59 PM.**
